Open source landscape ecology tools

### 2020-07-27 paper/paper.Rmd

<!-- elevatr + ?? - elevation -->
<!-- land cover -->
<!-- rgbif + ?? -->
<!-- climate data? -->
<!-- https://github.com/rspatial/geodata/blob/master/R/worldclim.R -->
<!-- rgee -->
<!-- look for rOpenSci packages -->

>>>>

A plethora of land cover products on global, continental, and country scales has been developed, however, most of them cannot be directly downloaded from R. 
The exception is the National Land Cover Database (NLCD) data for the USA, which can be accessed through the *FedData* package \cite{R-FedData}.
The *elevatr* package provides a synthesis raster elevation data based on the Amazon Web Services Terrain Tiles \cite{R-elevatr}.
It uses different elevation products for a given area and selected level of details (zoom/resolution).
<!-- how to cite (or maybe we should just add a link to?) - https://github.com/tilezen/joerd/blob/master/docs/data-sources.md -->
The *rgbif* gives access to data from the Global Biodiversity Information Facility (GBIF) portal \cite{R-rgbif}.
The `get_data()` function from the *raster* package allows to download climatic and bioclimatic data from WorldClim version 1.4, while the new in-development package *geodata*<!-- https://github.com/rspatial/geodata/blob/master/R/worldclim.R --> provides an access to WorldClim version 2 data.
Additionally, the *rgee package* \cite{R-rgee} gives access to an extensive catalog of data from Google Earth Engine, including climate data, land cover maps, and satellite imagery.
<!-- it is worth to search for more as there are a lot of changes going on -->

<!--JN: Max, do you want to add something to this list?-->
